---
title: "Practical Machine Learning Project"
author: "Ricardo Cecato Mavigno"
date: "24 de agosto de 2014"
output:
  html_document:
    number_sections: yes
---

Summary
================================================================================
As part of the Practical Machine Learning course from Johns Hopkins on Coursera, this work will try to predict the manner which subjects that where using devices that permit to collect dat about their activities did exercices. More information about the data can be found at this [link](http://groupware.les.inf.puc-rio.br/har).
We choose the Random Forest method to train the data, and the accuracy obtainded was:

- 100% for the in sample data.
- 99.3% for the out of sample data.

```{r, echo=TRUE}
#setup environment
library(caret)
library(randomForest)
setwd("~/Documents/R Scripts/PML")
```
Loading Data
================================

```{r, echo=TRUE,cache=TRUE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
              "pml-training.csv",method="curl")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
              "pml-testing.csv",method="curl")
pml_testing <- read.csv("data/pml-testing.csv",na.strings=c("NA",""))
pml_training <- read.csv("data/pml-training.csv",na.strings=c("NA",""))
#summary(pml_testing)
#summary(pml_training)
```

Cleaning Data
================================
```{r, echo=TRUE,cache=TRUE}
#keep only columns of interest
colNA <- sapply(pml_training, function(x) {sum(is.na(x))})
namesNotNA <- names(colNA[colNA==0])
pml_training2 <- subset(pml_training,select=namesNotNA)
colNA <- sapply(pml_testing, function(x) {sum(is.na(x))})
namesNotNA <- names(colNA[colNA==0])
pml_testing2 <- subset(pml_testing,select=namesNotNA)
pml_training2 <- pml_training2[,
                        -which(names(pml_training2) %in% c("X",
                                                "user_name",
                                                "raw_timestamp_part_1",
                                                "raw_timestamp_part_2",
                                                "cvtd_timestamp",
                                                "new_window",
                                                "num_window"))]
pml_testing2 <- pml_testing2[,
                        -which(names(pml_testing2) %in% c("X",
                                                "user_name",
                                                "raw_timestamp_part_1",
                                                "raw_timestamp_part_2",
                                                "cvtd_timestamp",
                                                "new_window",
                                                "num_window"))]

```
Split training data into training and test for cross validation
===============================================
```{r, echo=TRUE,cache=TRUE}
set.seed(289611)
inTrain <- createDataPartition(y=pml_training2$classe,p=0.6,list=FALSE)
training <- pml_training2[inTrain,]
testing <- pml_training2[-inTrain,]

```

Build Model using Random Forests
================================================
```{r, echo=TRUE,cache=TRUE}
model <- train(classe ~ .,data=training,
               method="rf",prox=TRUE,
               trControl=trainControl(method = "cv", number = 4, allowParallel = TRUE)) #build model using Random Forests and 4 fold train control
model
```

In sample accuracy
========================
```{r, echo=TRUE,cache=TRUE}
train.predict <- predict(model,training)
confusionMatrix(train.predict,training$classe)
```
The accuracy show by the confusionMatrix is 100% for the in sample predict.

Out of sample accuracy
======================
```{r, echo=TRUE,cache=TRUE}
test.predict <- predict(model,testing)
confusionMatrix(test.predict,testing$classe)
```
The accuracy show by the confusionMatrix is 99.3% for the out of sample predict.

Answering the Course Project submission
======================
```{r, echo=TRUE,cache=TRUE}
answers <- predict(model,pml_testing2)
answers <- as.character(answers)
answers

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE,
            col.names = FALSE)
    }
}

pml_write_files(answers)
```






